
---
title: "R Markdown - Assignment 3 in SDS 6103"
author: "John Andrew"
date: "2024-11-05"
output:
  html_document:
    toc: true
    toc_depth: 2
    code_folding: hide
  word_document:
    toc: true
    toc_depth: '2'
  pdf_document:
    toc: true
    toc_depth: 2
    number_sections: true
    latex_engine: xelatex
fontsize: 13pt
geometry: margin=1in
always_allow_html: true
---

---

# Illustration of R Markdown Capabilities

## Exploring the Use of R Markdown for Presentations, Notebooks, and Interactive Graphics

This document explores how R Markdown can be used to create presentations, notebooks, and interactive graphics, along with examples for each.

---

## R Markdown for Presentations

R Markdown supports creating various types of presentations:

1. **HTML Presentations**:
   - **ioslides**: Simple and clean HTML-based slides.
   - **Reveal.js**: Rich with themes, transitions, and interactivity.
   - **xaringan**: Highly customizable, built on `remark.js`.

2. **PDF Presentations**:
   - **Beamer**: LaTeX-based, suitable for academic and professional presentations.

3. **Microsoft PowerPoint**:
   - Directly generate `.pptx` files from R Markdown.

### Creating an HTML Presentation with ioslides

Example YAML header for ioslides:

```yaml
---
title: "Data Analytics Overview"
author: "John Andrew"
output: ioslides_presentation
---
```

This will create a basic HTML slide deck with ioslides.

---

### Creating a PDF Presentation with Beamer

Example YAML header for Beamer:

```yaml
---
title: "Statistical Analysis"
author: "John Andrew"
output: beamer_presentation
---
```

This setup will generate a Beamer PDF presentation with numbered sections.

---

### Creating a PowerPoint Presentation

Example YAML header for PowerPoint:

```yaml
---
title: "Business Intelligence"
author: "John Andrew"
output: powerpoint_presentation
---
```

This setup outputs directly to a `.pptx` file.

---

## Interactive Graphics with Shiny and HTML Widgets

R Markdown can produce interactive graphics using:
- **Shiny**: R's web framework for dynamic, reactive content.
- **HTML Widgets**: A collection of R packages for embedding interactive JavaScript widgets like `plotly` and `leaflet`.

### Interactive Plot Example (HTML Only)

To ensure compatibility with both HTML and PDF, the code is in HTML output.

```{r, eval = knitr::is_html_output()}
library(plotly)
plot_ly(data = mtcars, x = ~mpg, y = ~hp, type = 'scatter', mode = 'markers')
```

If i render as HTML, this chunk will display an interactive Plotly scatter plot.

---

### Interactive Map with Leaflet (HTML Only)

```{r, eval = knitr::is_html_output()}
library(leaflet)
leaflet() %>%
  addTiles() %>%
  addMarkers(lng = -122.4194, lat = 37.7749, popup = "San Francisco")
```

---

## R Notebooks in R Markdown

An R Notebook is an extension of R Markdown that supports inline execution and interactive analysis.

### Inline Plotting in R Notebook

```{r}
plot(mtcars$mpg, mtcars$hp, 
     main = "Miles per Gallon vs Horsepower",
     xlab = "Miles per Gallon", 
     ylab = "Horsepower")
```

### Using HTML Widgets in R Notebook (HTML Only)

```{r, eval = knitr::is_html_output()}
library(DT)
datatable(mtcars)
```

---

## Caching in R Notebooks

R Notebooks support caching, which saves results of computationally expensive code chunks.

```{r cache=TRUE}
Sys.sleep(5)  # Simulate a long computation
sum(rnorm(1e6))
```

---

## Parameterized Reports in R Notebooks

Example YAML for parameterized reports:

```yaml
---
title: "Parameterized Report"
output: html_notebook
params:
  sample_size: 20
  region: "West"
---
```

This feature allows flexible, dynamic reports based on parameter values.

---

## Using Other Language Engines in R Markdown

R Markdown supports different programming languages.

### Python Example (Illustrative Only)

```{python, eval=FALSE}
# Python calculation example
x = 10
y = 5
result = x * y
result
```

### SQL Example (Illustrative Only)

```{sql, eval=FALSE}
-- SQL query example
SELECT * FROM mtcars WHERE mpg > 20
```

### Bash Example (Illustrative Only)

```{bash, eval=FALSE}
# Bash command example
ls -l
```

---

